---
title: "Lecture06 R Functions"
author: "Michael Nguyen"
date: "10/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is H1 (Heading 1)
This is my work from lecture 6 for **BIMM143**

```{r}
# This is to demo a code chunk. This area is code, same as R envir't
plot(1:10)
```

## This is H2

### This is H3, etc. 




## Practice reading files (again...)
Here is to practice reading 3 different files

```{r}
read.table("Bimm143_L6_test1.txt", sep="," , header = TRUE)
```

```{r}
read.csv("Bimm143_L6_test1.txt")
```

```{r}
read.table("Bimm143_L6_test2.txt", sep="$", header = TRUE)
```

```{r}
read.table("Bimm143_L6_test3.txt", sep="", header=FALSE)
```


## Writing Functions 
### Use shortcut

```{r}
add <- function(x, y=1) {# sum the input x and y 
  x + y
  }
```


```{r}
add(1)

add(5,5)

# Erros: add(5, 5, 5) and add(5, "barry")
```


##Functions can be vectorized
```{r}
add(x=1, y=4) 
add(1, 4) 
add(1)
add( c(1, 2, 3) )
```


```{r}
add( c(1, 2, 3), 4 )
```



## Using Data frames and rescale funciton--> MAKE THINGS EASIER FOR YOURSELF
### You need a "name", "arguments", and "body"

```{r}
rescale <- function(x) {
rng <-range(x)
(x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale(1:10)
```



## Test, Fail, change test
### NA = Not Available

Testing some stuff

```{r}
rescale( c(1, 2, NA, 3, 10))
```


```{r}
x <- c(1,2,NA,3,10)
rng <- range(x, na.rm = TRUE)
rng
```



```{r}
rescale2 <- function(x) {
rng <-range(x, na.rm = TRUE)
(x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2( c(1,2,NA,3,10))
```
rescale2(x) also works because we have alread assigned that vector to x



##
##
##
##


```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) { 
  rng <-range(x, na.rm=na.rm)
  print("Hello")

  answer <- (x - rng[1]) / (rng[2] - rng[1])
  print("is it me you are looking for?")

  if(plot) {

    plot(answer, typ="b", lwd=4)
  }

  print("I can see it in ...") 
  return(answer)
}
```


```{r}
rescale3(1:10)
```
If insert return(answer) to after answer and before print, it will only show you hello



```{r}
plot(rescale3(x))
```



# Lectur 6 Section 2 Lab answers - MICHAEL NGUYEN

Install the **bio3d** package for sequence and structure analysis
```{r}
#install.packages("bio3d") <- don't download here in rmd. 
```


```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug 
s2 <- read.pdb("1AKE") # kinase no drug 
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA") 
s2.chainA <- trim.pdb(s2, chain="A", elety="CA") 
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b 
s2.b <- s2.chainA$atom$b 
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor") 
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor") 
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```


## Q1
The type of object returned from the read.pdf()function are protein expression levels depending if there is a drug or not. 


## Q2
Produce a new similar PDB object, containing a subset of atoms, from a given larger PDB object


##Q3 \
rug()


## Q4 better plot
Bar plot 


## Q5 
Dendrogram plot
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) ) 
plot(hc)
```
S1 and S3 both have drugs applied to it and therefore are more related. S1 isn't related because no drug. 
Bar plot or the plot above. 



## Q6
Name of new function: kinase

```{r}
kinase <- function(x) {
  # Vectorization of the read.pdb() function to reduce duplication. The function reads pdb coordinate data file, with the object being kinase proteins. 
  s1 <- read.pdb(x)  
  # Reduce duplicated ines of vectors for the trim.pdb() function. trim.pdb() produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object. This function is specifically looking at "hain "A" and Atom type (elety) "CA"
  s1.chainA <- trim.pdb(s1, chain="A", elety="CA") 
  # Reduced duplicaiton. From the smaller PDB object, code looks specifically at the B-factor of chain A. 
  s1.b <- s1.chainA$atom$b 
  # Reduced duplication for plot() function. plot() function plotted the protein data in a line graph. 
  plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
}
```



```{r}
# Function reads and plots following PDB data for the "4AKE" protein 
kinase("4AKE")

# Function reads and plots following PDB data for the "1AKE" protein 
kinase("1AKE")

# Function reads and plots following PDB data for the "1E4Y" protein 
kinase("1E4Y")
```

