---
title: "lec07 R Functions and Packages"
author: "Michael Nguyen"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions, revisited

Source my functions from last day

# RECAP

Control + Opt + I = Bring down R
```{r}
source("http://tinyurl.com/rescale-R")
```


take any numeric value and rescale it. 
```{r}
rescale(1:10)
```

```{r}
rescale( c(1, 10, 5, NA, 6))
```

na.rm = TRUE therfore we have the NA.


Errors:
rescale2(c(1:10, "barry"))
*Error: Input x should be numeric*

rescale(c(1:10, "barry"))
*Error in x - rng[1] : non-numeric argument to binary operator*



##Write new function for NA
Write a function to find where there are NA elements in two input vectors. 
1. Make some simple inputs where I **know** the answer
```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


Looked online and found the **is.na()**
```{r}
is.na(x)
```

and the **which** function tells me the elements/values are true
```{r}
which(is.na(x))
```

```{r}
is.na(y)
```

```{r}
which(is.na(y))
```


```{r}
is.na(x)
is.na(y)
```


The  **AND (&)** function requries two inputs that are TRUE to give a TRUE output
```{r}
is.na(x) & is.na(y)
```


Taking the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have
THIS IS MY WORKING SNIPPET!!
```{r}
sum(is.na(x) & is.na(y))
```

ex/of sum function
```{r}
sum( c(TRUE, TRUE, FALSE, TRUE))
```


##So let's turn it into a function (both_na)
### function_name <- **function**(input arguments) {body}
```{r}
both_na <- function (x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, y)
```

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```


```{r}
both_na(x, y1)
```
2 because two matching NAs. 

```{r}
both_na(x, y2)
```



```{r}
x2 <- c(NA, NA)
```


Test
```{r}
both_na(x2, y2)
```



RECYCLING VARIABLES/VALUES
example:
```{r}
plot(1:10, col=c("red", "blue", "green"))
```

```{r}
x <- c(NA, NA, NA)
x2 <- c(NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
x3 <- c(1, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x3, y3)
```

Viewing length
```{r}
length(x)
length(y3)
```




Adding stop() to new funciton **both_na2**. Add a check for when inputs x, y are *not* same length
```{r}
both_na2 <- function (x,y) {
  if(length(x) != length(y)) {
    stop("Inputs x and y should be the same length!")
  }
  sum(is.na(x) & is.na(y))
}
```

Error in "both_na2(x, y3)" : **Inputs x and y should be the same length!**



#Lab 7
## Write a function **grade** to determine an overall grade from a vector of student homework assignment scores dropping the lowest single slignment score. 

```{r}
#Student 1
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
which.min(s1)
s1[which.min(s1)]

#THIS
s1[-which.min(s1)]
#Student 2
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
which.min(s2)
s2[which.min(s2)]
```


Function, round 1
```{r}
grade <- function(s) {
  mean(s[-which.min(s)], na.rm = TRUE)
}
```

```{r}
grade(s1)
```

```{r}
grade(s2)
```

Is there an NA in the student's grade?
```{r}
is.na(s2)
any(is.na(s2))
any(is.na(s1))
```


```{r}
grade <- function(s) {
  if(any(is.na(s))) {
    warning("Student is missing a homework")
  }
  mean(s[-which.min(s)], na.rm = TRUE)
}
```

```{r}
grade(s1)
```


```{r}
grade(s2)
```


if you wanna really take out NA (if you only did one, and rest are NAs)
```{r}
grade <- function(s) {
  if(any(is.na(s))) {
    warning("Student is missing a homework")
  }
  sum(s[-which.min(s)], na.rm = TRUE)/(length(s) - 1)
}
```

```{r}
grade(s2)
```

```{r}
s3 <- c(100, NA, NA, NA, NA)
grade(s3)
```




```{r}
url <- "https://tinyurl.com/gradeinput"
hw <- read.csv(url, row.names=1)
```

```{r}
apply(hw, 1, grade)
```

