---
title: "Lec09 Unsupervised Learning Mini Proj"
author: "Michael Nguyen"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Exploratory data analysis
## Data input
```{r}
read.csv("WisconsinCancer.csv")
```
```{r}
# Save your input data file to a new 'data' directory
fna.data <- "data/WisconsinCancer.csv"
```

```{r}
wisc.df <-read.csv("WisconsinCancer.csv")
head(wisc.df)
```

"`r" means Rcode!

Note that the `id` and `diagnosis`cols will not be used for most of the folllowing steps

we have `r nrow(wisc.df)` samples in the data set
```{r}
# Find number of pts. 
nrow(wisc.df)
```


```{r}
table(wisc.df$diagnosis)
```

```{r}
# Converting the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
head(wisc.data)
```

Store the diagnosis for references in the future as a separate vector
```{r}
# Create diagnosis vector for later 
diagnosis <- wisc.df$diagnosis
```


Q1. How many observations are in this dataset?
- 569
nrow(wisc.df)

Q2. How many of the observations have a malignant diagnosis?
- 212
table(wisc.df)

Q3. How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.df)
```
```{r}
grep("_mean", colnames(wisc.df))
length(grep("_mean", colnames(wisc.df)))
```

These functions may be useful for answering the first 3 questions above.
dim() - gives ROWS x COLs, 
nrow() - gives the number of rows, 
table() - gives us table by cross-classifying factors,
length()- "size" of data you want to know 
grep() - pattern matching 


# 2. PCA
## Performing PCA
The next step in your analysis is to perform principal component analysis (PCA) on wisc.data.

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.
The input variables have significantly different variances.
Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before
```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```
THEY LOOK DIFFERENT... 
can try:
```{r}
round(colMeans(wisc.data), 3)
round(apply(wisc.data, 2, sd), 3)
```

MEANS and with SD are different! therefore `scale = TRUE`

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
```
==> Look at *Cumulative Proportion*
Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
- 0.4427

Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
- 3

Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?
- 7
```{r}
P <- summary(wisc.pr)
which(P$importance[3,] >0.9)[1]
# or `P$importance[row, col][1]
```


Has some structure. Useful to inspect. PC1 vs. PC2 = 60% 



## Interpreting PCA results
Now you will use some visualizations to better understand your PCA model. A common visualization for PCA results is the so-called biplot.

However, you will often run into some common challenges with using biplots on real-world data containing a non-trivial number of observations and variables. Here we will need to look at some alternative visualizations. You are encouraged to experiment with additional visualizations before moving on to the next section

Create a biplot of the wisc.pr using the biplot() function.

```{r}
plot(wisc.pr)
```

let's make a plot of PC1 vs. PC2 (with color)
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis, xlab = "PC1", ylab = "PC2")
```

Results = are like whoa. There is a split between cancer and malignant cancer diagnosis. 
```{r}
biplot(wisc.pr)
```

Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?
- It's messy and difficult to interpret... CRAP

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?
- It's somewhat the same as PC1 vs. PC2, but lower.
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3], col = diagnosis, xlab = "PC1", ylab = "PC3")
```


Because principal component 2 explains more variance in the original data than principal component 3, you can see that the **first plot has a cleaner cut separating** the two subgroups.

### Overall, the plots indicate that principal component 1 is capturing a separation of malignant from benign samples. This is an important and interesting result worthy of further exploration - as we will do in the next sections!




## Variance explained

In this exercise, you will produce scree plots showing the proportion of variance explained as the number of principal components increases. The data from PCA must be prepared for these plots, as there is not a built-in function in base R to create them directly from the PCA model.

As you look at these plots, ask yourself if there’s an ‘elbow’ in the amount of variance explained that might lead you to pick a natural number of principal components. If an obvious elbow does not exist, as is typical in some real-world datasets, consider how else you might determine the number of principal components to retain based on the scree plot.


```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```


## WILL DO LATER ^









# 3. Hierarchical Clustering 
## Hierarchical clustering of case data

The goal of this section is to do hierarchical clustering of the observations. Recall from our last class that this type of clustering does not assume in advance the number of natural groups that exist in the data.

As part of the preparation for hierarchical clustering, the distance between all pairs of observations are computed. Furthermore, there are different ways to link clusters together, with single, complete, and average being the most common linkage methods.

Scale the wisc.data data and assign the result to data.scaled.


```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```


```{r}
# Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.
data.dist <- dist(data.scaled)

# Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to wisc.hclust.
wisc.hclust <- hclust(d = data.dist, method = "complete")
```

## Results of hierarchical clustering
Let’s use the hierarchical clustering model you just created to determine a height (or distance between clusters) where a certain number of clusters exists.

Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

- h = 19
```{r}
plot(wisc.hclust)
abline(h = 19, col="red", lty=2)
```

## Selecting number of clusters
In this section, you will compare the outputs from your hierarchical clustering model to the actual diagnoses. Normally when performing unsupervised learning like this, a target variable (i.e. known answer or labels) isn’t available. We do have it with this dataset, however, so it can be used to check the performance of the clustering model.

When performing supervised learning - that is, when you’re trying to predict some target variable of interest and that target variable is available in the original data - using clustering to create new features may or may not improve the performance of the final model.

This exercise will help you determine if, in this case, hierarchical clustering provides a promising new feature.

Use cutree() to cut the tree so that it has 4 clusters. Assign the output to the variable wisc.hclust.clusters.

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
wisc.hclust.clusters
```

We can use the table() function to compare the cluster membership to the actual diagnoses.
```{r}
table(wisc.hclust.clusters, diagnosis)
```

Here we picked four clusters and see that cluster 1 largely corresponds to malignant cells (with diagnosis values of 1) whilst cluster 3 largely corresponds to benign cells (with diagnosis values of 0).

Before moving on, explore how different numbers of clusters affect the ability of the hierarchical clustering to separate the different diagnoses.

Q12. Can you find a better cluster vs diagnoses match with by cutting into a different number of clusters between 2 and 10?

- When you change the number clusters (k-value) it shows different number of clusters of different specificities. Too little of a k-value results in false posisitves while too high of a number results in better separation but too many spcifics can be confusing. 


# 4. OPTIONAL: K-means clustering
K-means clustering and comparing results
As you now know, there are two main types of clustering: hierarchical and k-means.

In this section, you will create a k-means clustering model on the Wisconsin breast cancer data and compare the results to the actual diagnoses and the results of your hierarchical clustering model. Take some time to see how each clustering model performs in terms of separating the two diagnoses and how the clustering models compare to each other.

Create a k-means model on wisc.data, assigning the result to wisc.km. Be sure to create 2 clusters, corresponding to the actual number of diagnosis. Also, remember to scale the data (with the scale() function and repeat the algorithm 20 times (by setting setting the value of the nstart argument appropriately). Running multiple times such as this will help to find a well performing model.

```{r}
wisc.km <- kmeans(wisc.data, centers= 2, nstart= 20)
wisc.km

# Use the table() function to compare the cluster membership of the k-means model (wisc.km$cluster) to the actual diagnoses contained in the diagnosis vector.

table(wisc.km$cluster, diagnosis)

#wisc.km$cluster vs. wisc.hclust.clusters
table(wisc.hclust.clusters, wisc.km$cluster)
```
Q13. How well does k-means separate the two diagnoses? How does it compare to your hclust results?
- Better as it groups them into clusters. 


# 5 Combining methods
Clustering on PCA results
In this final section, you will put together several steps you used earlier and, in doing so, you will experience some of the creativity and open endedness that is typical in unsupervised learning.

Recall from earlier sections that the PCA model required significantly fewer features to describe 70%, 80% and 95% of the variability of the data. In addition to normalizing data and potentially avoiding over-fitting, PCA also uncorrelates the variables, sometimes improving the performance of other modeling techniques.

Let’s see if PCA improves or degrades the performance of hierarchical clustering.

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.


```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2")
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
plot(hclust(dist(wisc.pr$x[,1:7]), method = "ward.D2"))
```


```{r}
table(grps, diagnosis)
# Sum Row 1 = 216!, same with Row 2 (353)
```

### In PC1: 28 False Positives!!! --> Misdiagnosised! Same with the PC2: 24 are false positives. 



```{r}
plot(wisc.pr$x[,1:2], col=grps)
plot(wisc.pr$x[,1:2], col=diagnosis)

# Note the color swap here as the hclust cluster 1 is mostly “M” and cluster 2 is mostly “B” as we saw from the results of calling table(grps, diagnosis). To match things up we can turn our groups into a factor and reorder the levels so cluster 2 comes first
```

```{r}
g <- as.factor(grps)
levels(g)
# Space
g <- relevel(g,2)
levels(g)
```

```{r}
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```


```{r eval=FALSE}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=grps)
rglwidget(width = 400, height = 400)
```


To include the interactive rgl plot in your HTML renderd lab report you can add the R code `rglwidget(width = 400, height = 400)` after you call the plot3d() function. It will look just like the plot above. Try rotating and zooming on this 3D plot.

Q14. How well does the newly created model with four clusters separate out the two diagnoses?
- It separates out them okay 

Q15. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.
- We got somoe false positives (healthy in cancerous cell populations and v.v.)



# 6. 









# 7. Prediction
We will use the `predict()` function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
Q17. Which of these new patients should we prioritize for follow up based on your results?
- Paitent number 2. RED IS MALIGNANT, BLACK IS BENIGN


# 8. PCA of protein sturcture data (Optional)
PCA of protein structure data
Visit the Bio3D-web PCA app http://bio3d.ucsd.edu and explore how PCA of large protein structure sets can provide considerable insight into major features and trends with clear biological mechanistic insight. Note that the final report generated from this app contains all the R code required to run the analysis yourself - including PCA and clustering. We will delve more into this type of analysis in the next class.

